cardinalR: Generating interesting high-dimensional data structures

### 2025-11-20 paper-cardinalR.tex
Linked shape generators

The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes, which can then be combined into a single dataset including scaling, rotation and translation. The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.

The main function, \texttt{gen\_multicluster()}, is an all-in-one function that includes generating individual shapes, handles scaling and rotating of these shapes, and combines the result into a single unified dataset. This function and associated workflow allows flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods. Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.

@@

The following is an example of a three shape multiclustered dataset. The first shape is gaussian, the second conical, and the third a cube.

@@

Here, the shapes have 200, 300 and 500 points respectively (\texttt{n}), are positioned in 4-D space according to a location matrix, \texttt{loc}, and stretched according to the \texttt{scale}. The details of the individual shape generators, and the noise elements are contained in the following sections.

>>>>

The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (\url{https://jayanilakshika.github.io/cardinalR/reference/index.html}), which can then be combined into a single dataset including scaling, rotation, and translation. The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.

The main function, \texttt{gen\_multicluster()}, is an all-in-one function that includes generating individual shapes, handling scaling and rotating of these shapes, and combining the result into a single unified dataset. This function and associated workflow allow flexible construction of complex, high-dimensional structures for evaluating clustering and dimension reduction methods. Figure \ref{fig:workflow} illustrates the workflow of \texttt{gen\_multicluster()}.

@@

The following is an example of a three-shape multiclustered dataset. The first shape is Gaussian, the second conical, and the third a cube.

@@

Here, the shapes have \(200\), \(300\), and \(500\) points respectively (\texttt{n}), are positioned in \(4\text{-}D\) space according to a location matrix, \texttt{loc}, and stretched according to the \texttt{scale}. The details of the individual shape generators and the noise elements are contained in the following sections.
